\vspace{-0.8in}Brief Announcement: A Family of Leaderless Generalized-Consensus Algorithms

### 2016-02-11 podc16/paper2.tex
fixes in the graph processing section, examples were missing
changes in the intro and minor other things

Many agreement algorithms inspired by Paxos and MultiPaxos~\cite{lamport2001paxos} rely on the existence of a unique leader replica which enforces an ordering on commands during fault-free periods.
In practice, the unique leader can easily become a performance bottleneck for the system.
Therefore, recently proposed agreement algorithms like Mencius~\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine}, EPaxos~\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments}, or Alvin~\cite{TurcuETAL14BeGeneralDontGiveUpConsistency} strive to increase performance by avoiding the use of a unique leader.
Mencius rotates the leader among the replicas but its structure remains close to MultiPaxos. 
However, EPaxos and Alvin are based on a novel algorithmic idea first introduced in EPaxos.

In this paper, we show that the core algorithmic idea underlying EPaxos can be used to build a family of leaderless algorithms parameterized by (1) a per-command dependency-set computation algorithm, a concept that we define in the paper, and (2) an algorithm ensuring agreement on a mapping from commands to dependency sets. %(2) an algorithm ensuring a per-command agreement on a dependency set and which is oblivious of any ordering between commands. 
An algorithm in the family is obtained by instantiating the two parameters to a concrete agreement algorithm and a concrete dependency-set computation algorithm, which are then used as sub-routines. 
Note that any generalized consensus algorithm can be used to instantiate parameter 2, but that only a weaker problem than generalized consensus needs to be solved because no ordering constraints need to be enforced.

@@

Our work sheds light on the common structure of algorithms like EPaxos and Alvin and allows to tune the performance of leaderless algorithms for specific scenarios by changing the underlying map-agreement and dependency-set algorithms.
We expect that these two tunable knobs will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by this new family of algorithms.

@@

We now present our abstract leaderless generalized consensus algorithm.
A process $p$ receiving a command $c$ proceeds in three phases. First, it calls a \textit{dependency-set} sub-algorithm (\Cref{dep-algo}) to determine a set of commands $cs$ that the command $c$ may depend on. Second, the process $p$ proposes the mapping  $c\mapsto cs$ to a \textit{map-agreement} sub-algorithm (\Cref{map-algo}); the map-agreement sub-algorithm may subsequently commit a mapping $c\mapsto cs'$ for $c$, in which case $cs'$ becomes the set of dependencies of $c$, and we say that $c$ is executable when its dependency set is empty or when all its dependencies are executable. 
Finally, when $c$ is executable, the process $p$ may execute $c$ by deciding the c-struct obtained by running the \emph{graph-processing} algorithm described in \Cref{graph-proc}.
Any other process can run the three phases independently of $p$ and eventually execute the command.
However, we show that a leaderless  generalized consensus algorithm is always safe and is live when at least a majority of the processes are correct.

@@

@@

\subsection{Implementing Leaderless Generalized Consensus}

Genearalized consensus~\cite{Lamport05GeneralizeConsensus} is a specification for state-machine replication algorithms. 
A generalized-consensus algorithm exposes the input actions $propose\left( c \right)$, for a command $c$, and $learn\left( \sigma \right)$, for a c-struct $\sigma$.
Moreover, a generalized-consensus algorithm must ensure that TODO.

We now propose an algorithm that uses the dependency-set computation abstraction and the map-agreement abstraction to implement generalized consensus. We therefore obtain a family of generalized consensus algorithms where each member of the family is obtained by using concrete dependency-set and map-agreement algorithms.

\todo{Explain the algorithm with pseudocode?}

Local variables: $m_p$, a map from commands to dependency sets, initially the empty map; $\sigma_p$, a c-struct, initially $\bot$.
\begin{itemize}[noitemsep,nolistsep]
    \item Upon ${\opn{propose}\left( c \right)}^p$: call ${\opn{announce}\left( c \right)}^p$.
    \item Upon ${\opn{suggest}\left( c,d \right)}^p$: call ${\opn{propose}\left( c,d \right)}^p$.
    \item Upon ${\opn{decide}\left( c,d \right)}^p$: insert the mapping $c \rightarrow d$ in the map $m_p$.
    \item Periodically, update $\sigma_p$ to $\opn{c-struct}\left( m_p \right)$, where $\opn{c-struct}\left( m_p \right)$ is defined below.
\end{itemize}

\subsection{Graph Processing}
\label{graph-proc}

@@

%\textbf{Orders induced by a graph.}
A directed graph $g$ \emph{induces a partial order} $\preceq_g$ on its vertices defined such that $c_1 \preceq_g c_2$ iff there is a path from $c_1$ to $c_2$ and there is no path from $c_2$ to $c_1$.
%Let $\opn{TOs}\left( g \right)$ be the set of total orders $o$ such that $\preceq_g \subseteq o$.
For example, $\preceq_h = \aset{\left( c_1,c_3 \right)}$. % and $\opn{TOs}\left( h \right)$ is the set of all permutations of $\aseq{c_1,c_2,c_3,c_4}$ in which $c_1$ comes before $c_3$.
If $c\in V\left( g \right)$, we write $\left[ c \right]_g$ for the strongly connected component of $g$ to which $c$ belongs. 
%\textbf{Equivalence relation induced by a graph.}
%A directed graph $g$ \emph{induces an equivalence relation} $\sim_g$ on its vertices defined such that $c_1 \sim_g c_2$ iff $c_1$ and $c_2$ belong to the same strongly connected component. We write $\left[ v \right]_g$ for the equivalence class of $v$ under $\sim_g$.
%\textbf{Linearizations of a dependency graph.}
We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every pair of vertices $c_1,c_2\in V\left( g \right)$, $c_1 \preceq_g c_2$ implies $c_1 \ll c_2$, and if $c_1 \sim_g c_2$ and $ c_1 <_{\left[ c_1 \right]_g} c_2$ hold then $c_1 \ll c_2$.
%\textbf{Partial order on graphs.}

@@

\subsection{Correctenss Proof}

>>>>

Many agreement algorithms inspired by Paxos and MultiPaxos~\cite{lamport2001paxos} rely on the existence of a unique leader process which enforces an ordering on commands during fault-free periods.
In practice, the unique leader can easily become a performance bottleneck for the system.
Therefore, recently proposed agreement algorithms like Mencius~\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine}, EPaxos~\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments}, or Alvin~\cite{TurcuETAL14BeGeneralDontGiveUpConsistency} strive to increase performance by avoiding the use of a unique leader.
Mencius rotates the leader among the processes but its structure remains close to MultiPaxos. 
However, EPaxos and Alvin are based on a novel algorithmic idea first introduced in EPaxos.

We show that the core idea underlying EPaxos can be captured in an abstract algorithm that uses two abstractions: (1), a per-command dependency-set computation algorithm, a concept that we define in the paper, and (2), an algorithm ensuring, for each command proposed to it, agreement on a dependency set. Both abstractions lend themselves well to implementations that avoid the use of a unique leader process.

Our abstract algorithm gives rise to a family of algorithms whose members are obtained by using a concrete map-agreement algorithm and a concrete dependency-set computation algorithm. 
Note that any generalized consensus algorithm can be used to implement (2), but only a weaker problem than generalized consensus needs to be solved because no ordering among proposal is required.

Our work has two main advantages: it enables modular correctness proofs of algorithms like EPaxos and it allows to tune the performance of implementations by changing the underlying map-agreement and dependency-set algorithms. We expect that these two tunable knobs will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by this new family of algorithms.

@@

@@

We now present our abstract leaderless generalized consensus algorithm.
A process $p$ receiving a command $c$ proceeds in three phases. First, it calls a \textit{dependency-set} sub-algorithm (\Cref{dep-algo}) to determine a set of commands $cs$ that the command $c$ may depend on. Second, the process $p$ proposes the mapping  $c\mapsto cs$ to a \textit{map-agreement} sub-algorithm (\Cref{map-algo}); the map-agreement sub-algorithm may subsequently commit a mapping $c\mapsto cs'$ for $c$, in which case $cs'$ becomes the set of dependencies of $c$, and we say that $c$ is executable when its dependency set is empty or when all its dependencies are executable. 
Finally, when $c$ is executable, the process $p$ may execute $c$ by deciding the c-struct obtained by running the \emph{graph-processing} algorithm described in \Cref{abstract-algo}.
Any other process can run the three phases independently of $p$ and eventually execute the command despite a crash of $p$.
However, we show that a leaderless  generalized consensus algorithm is always safe and is live when its underlying dependency-set and map-agreement algorithms are safe and live.

@@

\todo{Why not add the decide call to the interface, in order to guarantee liveness?}

@@

\subsection{The Abstract Generalized Consensus Algorithm}
\label{abstract-algo}
We now propose describe our abstract generalized consensus algorithm. 
Local variables: $m_p$, a map from commands to dependency sets, initially the empty map; $\sigma_p$, a c-struct, initially $\bot$.
\begin{itemize}[noitemsep,nolistsep]
    \item Upon ${\opn{propose}\left( c \right)}^p$: call ${\opn{announce}\left( c \right)}^p$.
    \item Upon ${\opn{suggest}\left( c,d \right)}^p$: call ${\opn{propose}\left( c,d \right)}^p$.
    \item Upon ${\opn{decide}\left( c,d \right)}^p$: insert the mapping $c \rightarrow d$ in the map $m_p$.
    \item Periodically, update $\sigma_p$ to $\opn{c-struct}\left( m_p \right)$, where $\opn{c-struct}\left( m_p \right)$ is defined below.
\end{itemize}

@@

%\textbf{Orders induced by a graph.}
A directed graph $g$ \emph{induces a partial order} $\preceq_g$ on its vertices defined such that $c_1 \preceq_g c_2$ iff there is a path from $c_1$ to $c_2$ and there is no path from $c_2$ to $c_1$.
%Let $\opn{TOs}\left( g \right)$ be the set of total orders $o$ such that $\preceq_g \subseteq o$.
For example, consider $h$ such that $V\left( h \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( h \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$. 
We have that $\preceq_h = \aset{\left( c_1,c_3 \right)}$. % and $\opn{TOs}\left( h \right)$ is the set of all permutations of $\aseq{c_1,c_2,c_3,c_4}$ in which $c_1$ comes before $c_3$.
If $c\in V\left( g \right)$, we write $\left[ c \right]_g$ for the strongly connected component of $g$ to which $c$ belongs. 
%\textbf{Equivalence relation induced by a graph.}
%A directed graph $g$ \emph{induces an equivalence relation} $\sim_g$ on its vertices defined such that $c_1 \sim_g c_2$ iff $c_1$ and $c_2$ belong to the same strongly connected component. We write $\left[ v \right]_g$ for the equivalence class of $v$ under $\sim_g$.
%\textbf{Linearizations of a dependency graph.}
We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every pair of vertices $c_1,c_2\in V\left( g \right)$, $c_1 \preceq_g c_2$ implies $c_1 \ll c_2$, and if $c_1 \sim_g c_2$ and $ c_1 <_{\left[ c_1 \right]_g} c_2$ hold then $c_1 \ll c_2$. For example, assuming that $c_2 \leq_{\aset{c_1,c_2}} c_1$, the linearizations of $h$ are $\aseq{c_2,c_1,c_3,c_4}$, $\aseq{c_2,c_1,c_4,c_3}$, $\aseq{c_2,c_4,c_1,c_3}$, and $\aseq{c_4,c_2,c_1,c_3}$.
%\textbf{Partial order on graphs.}

@@

\subsection{Correctness Proof}
